# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Navigate to the folder on your machine with your R code and the CSV.

Read the CSV into your program:

```{r}
setwd("/Users/loganfrederick/Projects/Coursera Johns Hopkins Data Science/Reproducible Research")
temp <- read.csv("RepData_PeerAssessment1/activity.csv")
```

## What is mean total number of steps taken per day?

Find the Daily Mean, Median, and Total Number of Steps while ignoring missing values.

Provide a Histogram of the daily steps.

```{r}
dailyStepsMean <- tapply(temp$steps,temp$date,mean,na.rm=TRUE)
dailyStepsMedian <- tapply(temp$steps,temp$date,median,na.rm=TRUE)
dailyStepSum <- tapply(temp$steps,temp$date,sum,na.rm=TRUE)

#Daily Means
print(dailyStepsMean)

#Daily Medians
print(dailyStepsMedian)

#Histogram
hist(dailyStepSum,breaks=15)

```

## What is the average daily activity pattern?

A Time Series plot of the average number of steps taken at five minute intervals.

The interval with the highest number of steps is #835.

```{r}
library(ggplot2)

stepsPerInterval <- data.frame(tapply(temp$steps, temp$interval, mean, na.rm=TRUE))
names(stepsPerInterval) <- "steps"
stepsPerInterval$interval <- as.integer(row.names(stepsPerInterval))
ggplot(stepsPerInterval) + aes(x=interval, y=steps) + geom_line()
which.max(stepsPerInterval$steps)
```

## Imputing missing values

The number of missing "steps" values is 2304:
```{r}
sum(is.na(temp$steps))
```

```{r}
#Replace NAs with a mean from a random day
temp[is.na(temp)] <- sample(as.vector(dailyStepsMean), 1)

#New daily means and medians
newDailyStepsMean <- tapply(temp$steps,temp$date,mean,na.rm=TRUE)
newDailyStepsMedian <- tapply(temp$steps,temp$date,median,na.rm=TRUE)
newDailyStepSum <- tapply(temp$steps,temp$date,sum,na.rm=TRUE)

#New Daily Means
print(newDailyStepsMean)

#New Daily Medians
print(newDailyStepsMedian)

hist(newDailyStepSum,breaks=15)
```


## Are there differences in activity patterns between weekdays and weekends?

Yes, comparing the lattice plots between weekends and weekdays shows at a glance clear difference in activity.

```{r}
temp$date <- as.Date(temp$date,"%Y-%m-%d")

avgStepInt <- aggregate(temp$steps,temp[3],mean,na.rm=TRUE)
merged_temp <- merge(x=temp,y=avgStepInt,by="interval",all.x=TRUE)
merged_temp$steps <- ifelse(is.na(merged_temp$steps),round(merged_temp$x,digits=2),merged_temp$steps)

factor <- merged_temp[,-c(4)]
factor$flag <- as.factor(ifelse(weekdays(factor$date) %in% c("Saturday","Sunday"),"Weekend","Weekdays"))
head(factor)

plot_temp <- aggregate(factor$steps,factor[c(1,4)],mean)

library(lattice)
xyplot(x ~ interval | flag,data = plot_temp,layout=c(1,2),type = "l",xlab = "Interval",ylab = "Number of steps")
```